Formalizing Stack Safety as a Security Property

### 2021-09-30 submission4/main.tex
Getting closer to 12 pages :-)

What does ``stack safety'' mean, precisely? The phrase is associated with a
variety of compiler,
run-time, and hardware mechanisms for protecting stack
memory, but these mechanisms typically lack precise specifications,
relying instead on informal descriptions and examples of the bad
behaviors that they prevent.

@@

Labels under the boxes indicate an assignment of the address to a
{\em domain}, which is a particular set of intended access
rights. Unless otherwise marked, each address retains its domain as
sequential instructions execute. %, as hinted by their colors.
%\bcp{where do we see this happening?  Is it important?}.
Domains are not part of the machine state; rather, they are part
of the {\em context} of a state that describes the security restrictions
that the enforcement mechanism must adhere to.
(We will introduce some other parts of the context as needed.)
The possible domains
are \(\unsealed\) (in the stack and accessible to writes), \(\sealed{\depth}\)
(in the stack and reserved for the caller at depth \(\depth\)), or \(\outside\)
(outside the stack entirely.)
% These are abbreviated \(U\) and \(S(\dots)\)\bcp{Nope} in the figures.
Global variables and code, which are not shown in the figures, would
also be labeled \(\outside\). Return pointers, labeled ``RetPtr,''
are shaded black and ignored for simplicity for now---they will
be treated properly later.

@@

\paragraph*{Machine model}
Our properties are defined in terms of an abstract machine model that makes very few
assumptions about the concrete machine that instantiates it. We require a
register machine with a program counter (\(\PCname\)) and there must be a well-defined
mechanism for determining the bounds of a stack frame, and for recognizing when a call
has returned. In this paper we discuss machines with a stack pointer (\(\rsp\)),
in which the current stack frame is everything below the stack pointer and above the
previous frame, and a call from instruction \(\addr_\PCname\) with stack pointer \(\addr_\rsp\)
is considered to have returned when the pc is at \(\addr_\PCname + 4\) and the stack pointer
is again at \(\addr_\rsp\). But our properties apply equally to machines with
other architectural conventions, such as those involving frame pointers.
A machine may use its non-stack memory in any way, including
as a heap, as long as the stack has static bounds. It might also possess additional
state beyond its registers and memory addresses, such as the state of a hardware
enforcement mechanism.

@@

\paragraph*{What is a ``call''?}  One significant challenge in formulating our
properties arises from the fact that we want to define stack safety at the
machine-code level for a variety of ISAs.  There may be no single ``call''
instruction that does all the
required manipulation of the \(\PCname\) and stack pointer. Instead, there
might be just a simple branch instruction that stores the old \(\PCname\) in
a register, with the rest of the call sequence performed by other
instructions either before or after the jump.  Similarly, there may be no
``return'' instruction, but rather a ``return sequence'' involving an
indirect branch at some point.  Thus, the same instruction opcode may be
used in some places as part of a call or return sequence and in other places
for other purposes.

@@

In this section we explore how an existing enforcement mechanism called {\em
  Depth Isolation}~\cite{DBLP:conf/sp/RoesslerD18} implements
the formal stack-safety property described in the previous
section, and why its optimized form, {\em Lazy Tagging and Clearing}, does not.
We validate this claim using property-based random testing in \cref{sec:testing}.

>>>>

What does ``stack safety'' mean? The phrase is associated with a
variety of compiler,
run-time, and hardware mechanisms for protecting stack
memory, but these mechanisms typically lack precise specifications,
relying instead on informal descriptions and examples of the bad
behaviors that they prevent.

@@

Labels under the boxes indicate an assignment of the address to a
{\em domain}---a set of intended access
rights. Unless otherwise marked, each address retains its domain as
instructions execute. %, as hinted by their colors.
%\bcp{where do we see this happening?  Is it important?}.
Domains are not part of the machine state; rather, they are part
of the {\em context} of a state that describes the security restrictions
that the enforcement mechanism must adhere to.
The possible domains
are \(\unsealed\) (in the stack and accessible to writes), \(\sealed{\depth}\)
(in the stack and reserved for the caller at depth \(\depth\)), or \(\outside\)
(outside the stack entirely.)
% These are abbreviated \(U\) and \(S(\dots)\)\bcp{Nope} in the figures.
Global variables and code, which are not shown in the figures, would
also be labeled \(\outside\). Return pointers, labeled ``RetPtr,''
are shaded black and ignored for simplicity for now---they will
be treated properly later.
(We will introduce some other parts of the context as needed.)

@@

\paragraph*{Machine model}
Our properties are defined in terms of an abstract model that makes few
assumptions about the concrete machine that instantiates it. We assume a
register machine with a program counter (\(\PCname\)) and well-defined
mechanisms for determining the bounds of a stack frame and recognizing when a call
has returned. Our examples consider machines with a stack pointer (\(\rsp\)),
in which the current stack frame is everything below the stack pointer and above the
previous frame, and where a call from instruction \(\addr_\PCname\) with
stack pointer \(\addr_\rsp\)
is considered to have returned when the pc is at \(\addr_\PCname + 4\) and the stack pointer
is again at \(\addr_\rsp\). However, our properties apply equally to
other architectural conventions, such as frame pointers.
The non-stack memory can be used in any way, including
as a heap, as long as the stack has static bounds. There can be
state other than registers and memory, such as the internal state of a hardware
enforcement mechanism.

@@

\paragraph*{What is a ``call''?}  One significant challenge in formulating our
properties arises from the fact that we want to define stack safety at the
machine-code level for a variety of ISAs.  There may be no single ``call''
instruction that does all the
required manipulation of the \(\PCname\) and stack pointer. Instead, the
same instruction opcode may be
used in some places in a call or return sequence and in other places
for other purposes.
\ifspace
For example, there
might be just a simple branch instruction that stores the old \(\PCname\) in
a register, with the rest of the call sequence performed by other
instructions either before or after the jump.  Similarly, instead of a single
``return'' instruction,  there may be a ``return sequence'' involving an
indirect branch at some point.
\fi

@@

We next explore how an \ifspace existing \fi enforcement mechanism called {\em
  Depth Isolation}~\cite{DBLP:conf/sp/RoesslerD18} implements
the formal stack-safety property described in the previous
section, and why its optimized form, {\em Lazy Tagging and Clearing}, does not.
We validate this claim using property-based random testing in \cref{sec:testing}.
